Towards practical dynamic induction control of wind farms: analysis of optimally controlled wind-farm boundary layers and sinusoidal induction control of first-row turbines

### 2018-05-23 revision/reply_to_revs_WM2.tex
R1C1 finalized

\textbf{Response:} This is an interesting comment. It is correct that this contrast results from the way turbines and their controls are modeled in the current study and that of Ciri et al. The reference disk-based thrust coefficient $C_T' = 2$ used here actually implies intrinsic self-optimization of blade pitch and generator torque to local flow conditions, even though these actions are not resolved by our current actuator disk formulation which directly controls $C_T'$. In case these degrees of freedom would be resolved, for instance using an actuator line model as in Ciri et al., the ESC from the latter study could e.g. be used to optimize torque controller gain in order to achieve an effective $C_T' \approx 2$.

We have included the fact that the reference case already implies self-optimization in the revised manuscript as follows (p 4, l. 21):

``
A conventionally (greedily) controlled wind farm with steady $C_T' = 2$ was defined as a reference case. \textbf{Note that this would correspond to a farm with ideal turbines for which generator torque is being controlled dynamically to track the maximum power point at the Betz limit perfectly.} Several different optimal control cases were defined, ...
''

However, even given the discussion above, it is not unimaginable that a turbine could temporarily outperform this greedy control setting, e.g. by timely reacting to incoming flow accelerations. This is what we would refer to as self-optimization. {\color{blue} We have explicitly defined this in the text as follows (p ??):

``
...This suggests that self-optimization \textbf{(i.e. temporarily outperforming conventional greedy control in response to incoming flow events)} is very limited: the optimal controls for a given turbine are designed to create ...
'' }

(IS DIT OKE?? Kan het deel in blauw misschien beter gewoon weglaten?)

@@

\textbf{Response:} Indeed, this is correct. We have mentioned this explicitly in the revised manuscript as follows (p. ??, l. ??):

@@

We have slightly modified the statement in the revised manuscript to avoid any confusion with regard to possibly having the chord length as a control variable. (p. ??, l. ??):

``
The resulting maximum thrust coefficient of 3.5 can, e.g., be \textbf{attained by the NREL 5MW turbine with a slight modification in the rotor design}, i.e. a 50\% increase in blade chord length and an operational tip speed ratio 25\% higher than 
the original design value (see Goit and Meyers, 2015, Appendix A). \textbf{Furthermore, given such redesign, dynamic reductions from this value could be realized through blade pitch control, for which actuation in the order of $10^\circ/$s is possible (see, e.g., Jonkman et al. 2009).} 
''

@@

\textbf{Response: } Indeed, we seem to have overlooked specifying the exact turbine dimensions in Section 2.2. We have updated the manuscript as follows (l XX p YY):

@@

Both concepts described above are discussed in more detail in the MM17 paper. To keep the description of the case setup concise in the current work, we simply add that $T = 240$ s corresponds to the time to pass four turbine rows, and further refer to MM17 for further elaboration of methodological choices. We have updated the revised manuscript as follows (l XX p XX):

@@

and (l XX p XX, end of Section 2.2):

@@

We have added the requested TI values in the discussion around Figure 4 as follows (p ?? l ??):

@@

We have updated the manuscript to be more explicit in the sense that results are not overall conclusions for wind-farm control in general, but should be interpreted as observations of the current optimal control case. See the beginning of Section 3 (l XX p XX):

@@

We have incorporated this in the revised manuscript as follows (p ??):

@@

and (p ??):

@@

\begin{figure}
	\includegraphics[width=\textwidth]{upstream_vel.eps}
	\caption{Disk-averaged axial velocity upstream of the first-row turbines for column 1 and column 2 in Figure 2 of the manuscript. Vertical dotted lines demarcate the optimization window boundaries (i.e. at integer multiples of $T_A$) \label{fig:upstream_vel}}
\end{figure}

@@

\textbf{Response: } We thank the reviewer for this suggestion. Dynamic reductions from the maximum $C_T' = 3.5$ could, for instance, be achieved by using the fast pitch actuators with which modern turbines are equipped. We have added this comment to the discussion as follows (p ??):

``
The resulting maximum thrust coefficient of 3.5 can, e.g., be \textbf{attained by the NREL 5MW turbine with a slight modification in the rotor design}, i.e. a 50\% increase in blade chord length and an operational tip speed ratio 25\% higher than 
the original design value (see Goit and Meyers, 2015, Appendix A). \textbf{Furthermore, given such redesign, dynamic reductions from this value could be realized through blade pitch control, for which actuation in the order of $10^\circ/$s is possible (see, e.g., Jonkman et al. 2009).} 
''

@@

We have incorporated this discussion in the revised version of the manuscript as follows (p ?? l ??):

@@

\textbf{Response: } We thank the reviewer for this comment. Agreed, the downstream turbines could also be subjected to increased fatigue loading, and we have included this in the revised manuscript as follows (p ??, l ??):

@@

>>>>

\textbf{Response:} This is an interesting comment. It is correct that the difference originates from the way turbines and their controls are modeled in the current study compared to that of Ciri et al. The reference disk-based thrust coefficient $C_T' = 2$ used here actually implies intrinsic self-optimization of blade pitch and generator torque to local flow conditions, even though these actions are not resolved by our current actuator disk formulation which directly controls $C_T'$. In case these degrees of freedom would be resolved, for instance using an actuator line model as in Ciri et al., the ESC from the latter study could e.g. be used to optimize torque controller gain in order to achieve an effective $C_T' \approx 2$.

We have included the fact that the reference case already implies self-optimization in the revised manuscript as follows (p 4, line 21):

``
A conventionally (greedily) controlled wind farm with steady $C_T' = 2$ was defined as a reference case. \textbf{Note that this would correspond to a farm with ideal turbines for which generator torque is being controlled dynamically to track the maximum power point at the Betz limit perfectly. In a real turbine controller this may, e.g., be achieved with the ESC proposed by Ciri et al. (2017).} Several different optimal control cases were defined, ...
''

@@

\textbf{Response:} Indeed, this is correct. We have mentioned this explicitly in the revised manuscript as follows (p. 4, line 25):

@@

We have slightly modified the statement in the revised manuscript to avoid any confusion with regard to possibly having the chord length as a control variable. (p. 17, line 29):

``
The resulting maximum thrust coefficient of 3.5 can, e.g., be \textbf{attained by the NREL 5MW turbine with a slight modification in the rotor design}, i.e. a 50\% increase in blade chord length and an operational tip speed ratio 25\% higher than 
the original design value (see Goit and Meyers, 2015, Appendix A). \textbf{Furthermore, given such redesign, dynamic reductions from this value could be realized through blade pitch control, for which actuation rates in the order of $10^\circ/$s are possible (see, e.g., Jonkman et al. 2009).} 
''

@@

\textbf{Response: } Indeed, we seem to have overlooked specifying the exact turbine dimensions in Section 2.2. We have updated the manuscript as follows (p. 4, line 15):

@@

Both concepts described above are discussed in more detail in the MM17 paper. To keep the description of the case setup concise in the current work, we simply add that $T = 240$ s corresponds to the time to pass four turbine rows, and further refer to MM17 for further elaboration of methodological choices. We have updated the revised manuscript as follows (p. 4, line 20):

@@

and (p. 4, line 27):

@@

We have added the requested TI values in the discussion around Figure 4 as follows (p. 6, line 19):

@@

We have updated the manuscript to be more explicit in the sense that results are not overall conclusions for wind-farm control in general, but should be interpreted as observations of the current optimal control case. See the beginning of Section 3 (p. 8, line 5):

@@

We have incorporated this in the revised manuscript as follows (p. 14, line 7):

@@

and (p. 14, line 14):

@@

\begin{figure}
	\centering
	\includegraphics[width=0.65\textwidth]{upstream_vel.eps}
	\caption{Disk-averaged axial velocity upstream of the first-row turbines for column 1 and column 2 in Figure 2 of the manuscript. Vertical dotted lines demarcate the optimization window boundaries (i.e. at integer multiples of $T_A$) \label{fig:upstream_vel}}
\end{figure}

@@

\textbf{Response: } We thank the reviewer for this suggestion. Dynamic reductions from the maximum $C_T' = 3.5$ could, for instance, be achieved by using the fast pitch actuators with which modern turbines are equipped. We have added this comment to the discussion as follows (p. 17, line 29):

``
The resulting maximum thrust coefficient of 3.5 can, e.g., be \textbf{attained by the NREL 5MW turbine with a slight modification in the rotor design}, i.e. a 50\% increase in blade chord length and an operational tip speed ratio 25\% higher than 
the original design value (see Goit and Meyers, 2015, Appendix A). \textbf{Furthermore, given such redesign, dynamic reductions from this value could be realized through blade pitch control, for which actuation rates in the order of $10^\circ/$s are possible (see, e.g., Jonkman et al. 2009).} 
''

@@

We have incorporated this discussion in the revised version of the manuscript as follows (p. 23, line 11):

@@

\textbf{Response: } We thank the reviewer for this comment. Agreed, the downstream turbines could also be subjected to increased fatigue loading, and we have included this in the revised manuscript as follows (p. 24, line 7):

@@

\clearpage

### 2018-05-23 revision/revised_WM02.tex
R1C1 finalized

A conventionally (greedily) controlled wind farm with steady $C_T' = 2$ was defined as a reference case. \revision{Note that this would correspond to a farm with ideal turbines for which generator torque is being controlled dynamically to track the maximum power point at the Betz limit perfectly.} Several different optimal control cases were defined, based on the wind-turbine response time $\tau = $ 0, 5, or 30 s (instantaneous, fast, or slow response, see Eq. \ref{eq:ctihat_constraint}) and the maximal thrust coefficient $C_{T,\text{max}}' = 2$ or~3 \revision{, with thrust forces that can respectively only be reduced (underinductive), or also increased (overinductive) compared to the Betz optimum at $C_T' = 2$} (see Eq. \ref{eq:boxct_constraint}). Cases are denoted as C<X>t<Y>, where X and Y represent $C_{T,\text{max}}'$ and $\tau$ respectively, e.g. C3t30  for the case with $C_{T,\text{max}}' = 3$ and $\tau = 30$ s. \revision{The choice of (and sensitivity to) setup parameters is further elaborated in MM17.}

@@

Figure~\ref{fig:multirow}b shows results from simulations in which the controls are applied \emph{for all rows up to a certain row}, i.e. R1--R3 indicates the application of optimized controls generated by case C3t5 to rows 1, 2, and 3. An interesting observation from this figure is that , for any row $i$ except the last one, the power potential as observed in case C3t5 is almost fully recovered by only applying the optimal controls up to row $i-1$. This suggests that self-optimization \revision{(i.e. temporarily outperforming conventional greedy control in response to incoming flow events)} is very limited: the optimal controls for a given turbine are designed to create favorable flow conditions in the downstream rows instead of increasing local power. Furthermore, although the discussion in the previous paragraph has shown that downstream controls are optimized with the upstream actions in mind, the converse is not true: upstream control actions do not require a specific downstream response in order to increase power in that downstream row.

>>>>

A conventionally (greedily) controlled wind farm with steady $C_T' = 2$ was defined as a reference case. \revision{Note that this would correspond to a farm with ideal turbines for which generator torque is being controlled dynamically to track the maximum power point at the Betz limit perfectly. In a real turbine controller this may, e.g., be achieved with the ESC proposed by \cite{ciri2017large}.} Several different optimal control cases were defined, based on the wind-turbine response time $\tau = $ 0, 5, or 30 s (instantaneous, fast, or slow response, see Eq. \ref{eq:ctihat_constraint}) and the maximal thrust coefficient $C_{T,\text{max}}' = 2$ or~3 \revision{, with thrust forces that can respectively only be reduced (underinductive), or also increased (overinductive) compared to the Betz optimum at $C_T' = 2$} (see Eq. \ref{eq:boxct_constraint}). Cases are denoted as C<X>t<Y>, where X and Y represent $C_{T,\text{max}}'$ and $\tau$ respectively, e.g. C3t30  for the case with $C_{T,\text{max}}' = 3$ and $\tau = 30$ s. \revision{The choice of (and sensitivity to) setup parameters is further elaborated in MM17.}

@@

Figure~\ref{fig:multirow}b shows results from simulations in which the controls are applied \emph{for all rows up to a certain row}, i.e. R1--R3 indicates the application of optimized controls generated by case C3t5 to rows 1, 2, and 3. An interesting observation from this figure is that , for any row $i$ except the last one, the power potential as observed in case C3t5 is almost fully recovered by only applying the optimal controls up to row $i-1$. This suggests that self-optimization is very limited: the optimal controls for a given turbine are designed to create favorable flow conditions in the downstream rows instead of increasing local power. Furthermore, although the discussion in the previous paragraph has shown that downstream controls are optimized with the upstream actions in mind, the converse is not true: upstream control actions do not require a specific downstream response in order to increase power in that downstream row.
